Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-07-16 snapshot_system/cds.tex

\section{CDS Analysis}
[Purpose of using CDS]
CDS is a small percentage of the global data index that we into DHT as a shared index. 
With CDS as as a public data resource, each VM's snapshot backup has no need to 
store its own copies for data that already exist in CDS, instead they can just
store a reference. 
We expect CDS to contain most of the data related to operating system, software and libraries.

[What's the benifits]
Using CDS as a common data source brings us several benefits: 
First, the data we collect into CDS are commonly used in many VMs, 
thus the CDS index, a special subset of global index, has very high deduplication efficency.
Second, in our architecture the CDS data are the only data that can be shared between different VMs,
by carefully control the replication degree of CDS data, we could improve the fault-tolerancy
in such a data sharing environment. Finally, the CDS index is small enough to be put into a 
in-memory DHT, thus it avoids the excess memory usage of global index 
while brings us the majority of deduplication efficiency of it.

[Describe how CDS is generated]

[Explain why CDS scales effectively as data grow]
Our empirical study based on VM images from production environment\cite{ieeecloud} showed that the 
frequency of data duplication follows Zipf-like distribution\cite{zipf}, 
with the exponent $\alpha$ between 0.65 ~ 0.70. 
As a result, it can be proved that deduplication efficiency of CDS index is scalable:

[Describe the dedup efficiency model in detail]
Let $S_c$ be the total number of FPs in CDS, $S_g$ be the total number of FPs in global index, $F$ be the size of a FP, $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by CDS, $D$ be the amount of data on each node, and $B$ be the average block size. The deduplication efficiency will be:
\begin{equation}
  E = (S_c / S_g)^{1-\alpha}
\end{equation}
The $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, S_g = N*D/B
\end{equation}
By replacing $S_c$ and $S_g$ in the first formula, we have:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}

Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of CDS is only controlled by the its memory usage.

>>>>

\section{PDS Analysis}
[Purpose of using PDS]
PDS is a small percentage of the global data index that we into DHT as a shared index. 
With PDS as as a public data resource, each VM's snapshot backup has no need to 
store its own copies for data that already exist in PDS, instead they can just
store a reference. 
We expect PDS to contain most of the data related to operating system, software and libraries.

[What's the benifits]
Using PDS as a common data source brings us several benefits: 
First, the data we collect into PDS are commonly used in many VMs, 
thus the PDS index, a special subset of global index, has very high deduplication efficency.
Second, in our architecture the PDS data are the only data that can be shared between different VMs,
by carefully control the replication degree of PDS data, we could improve the fault-tolerancy
in such a data sharing environment. Finally, the PDS index is small enough to be put into a 
in-memory DHT, thus it avoids the excess memory usage of global index 
while brings us the majority of deduplication efficiency of it.

[Describe how PDS is generated]

[Explain why PDS scales effectively as data grow]
Our empirical study based on VM images from production environment\cite{ieeecloud} showed that the 
frequency of data duplication follows Zipf-like distribution\cite{zipf}, 
with the exponent $\alpha$ between 0.65 ~ 0.70. 
As a result, it can be proved that deduplication efficiency of PDS index is scalable:

[Describe the dedup efficiency model in detail]
Let $S_c$ be the total number of FPs in PDS, $S_g$ be the total number of FPs in global index, $F$ be the size of a FP, $N$ be the number of nodes in the cluster, $m$ be the memory on each node that are used by PDS, $D$ be the amount of data on each node, and $B$ be the average block size. The deduplication efficiency will be:
\begin{equation}
  E = (S_c / S_g)^{1-\alpha}
\end{equation}
The $S_c$ and $S_g$ can be expressed as:
\begin{equation}
S_c = N*m/F, S_g = N*D/B
\end{equation}
By replacing $S_c$ and $S_g$ in the first formula, we have:
\begin{equation}
  E = (\frac{m*B}{F*D})^{1-\alpha}
\end{equation}

Since $B$, $D$ and $F$ are pre-configured constants, the deduplication efficiency of PDS is only controlled by the its memory usage.

### 2013-07-16 snapshot_system/design.tex

Overall, our contributions are:
\begin{enumerate}
\item {\em CDS. parallelism, localized}
We designed a highly efficient global deduplication scheme using common data between VMs as the primary
deduplication target, which adopts to our resource-constrained production environment very well. We developed a fast
approxmation algorithm to extract the commonly used data from VM snapshot backups.

@@

\item {\em Fault-tolerant.}
Golbal deduplication brings data sharing across different users VM snapshots, therefore a single failure
on some data can break many VMs' snapshots. We avoid such all-to-all data dependency by restricting
the cross VM data sharing with the CDS. By taking extra care of protecting CDS data, 
the overall degree of fault tolerancy is guaranteed.

@@

% Figure~\ref{fig:arch} shows the architecture view of our snapshot service
% at each node. The snapshot broker provides the functional interface for  snapshot backup, access, and deletion.
% The inner-VM  deduplication is conducted by the broker to access meta data in the snapshot data store
% and we discuss this in details in Section~\ref{sect:innerVM}.
% The cross-VM deduplication is conducted by the broker to access 
% a common data set (CDS) (will discuss in Section~\ref{sect:crossVM},
% whose block hash index is stored in a distributed memory cache. 
% \begin{figure}[htbp]
%   \centering
%   %\epsfig{file=images/arch.eps, height=2in, width=2.66in}
%   \epsfig{file=images/arch.eps, width=3.9in}
%   \caption{Snapshot backup architecture of each node.}
%   \label{fig:arch}
% \end{figure}
% The snapshot store 
% supports data access operations such as \emph{get}, \emph{put} and \emph{delete}.
% Other operations include data block traverse and resource usage report.
% The snapshot data does not need to be
% co-located with VM instances, and in fact they can even live in a different cluster to improve the 
% data reliability: when one cluster is not available, we are still able to restore its VMs from another cluster which
% holds its snapshot data. 
% %to reduce the impact to our users.
% %Get interface accepts a piece of data, write it to the underline data file, and return
% %a reference to the caller. This reference then can be used in the put interface to
% %retrive or delete the data, thus the caller of put interface must preserve the
% %data reference for future use. 
% %In addition to above standand data access operations, the snapshot service also supports
% %\emph{scan} and \emph{quota} methods. Scan allows us to traverse all the data blocks of each VM
% %in the snapshot store, and quota is used to acknowledge user how much space he
% %has actually used.

@@

% \subsection{Block Storage Device}
\subsection{CDS}
Although locality based data reduction can remove most of the inner-VM duplications,
it's not able to solve the data duplication cross VMs. Different VMs still share large amount
of common data such that their snapshot backups would have a lot of data in common.
Our observations on Aliyun's real VM user data reveals several major sources of cross-VM data duplication:
  \begin{enumerate}
  \item \emph{System-related data}: These data are generated by public third-parties, they are copied/downloaded/installed into VM disks either automatically or manually. Once installed, operations on such data are mostly read only until software updates arrive. For example, data of operating systems, some widely-used software such as Apache and MySQL, and their documations all fall into this category.
  \item \emph{User-generated data}: These data are generated by individual user's behavior, either directly or indirectly. They are much less duplicated than the system-related data, but the zipf-like distribution indicates that a small amount of data in this category could represent most of data duplications.
  \item \emph{Zero-filled data}: Like previous studies [refs here], we've observed that zero-filled data exist pervasively at system wide. They are almost like the spaces and commas in text articles. Under content-based chunking, they were distilled as zero-filled blocks with the maximum allowed length. 
%Zero-filled blocks are mostly generated by user behavior to preserve the space for future data, thus we treat them as a special case of user-generated data.
  \end{enumerate}

Without eliminating the data duplication between VM snapshot backups, storage space is severely wasted when the number of VMs increases.
To address this problem, we developed a technique called \emph{Common Data Set (CDS)} to eliminate data duplication for all three situations above. 
CDS is a public data library that shared by all VM snapshot backups in the cluster, 
which consists of the most popular data blocks in VM snapshot backups. It is generated, 
managed and accessed in an uniform manner by all VMs such that [write some fancy system characteristic stuff here].

\subsection{CDS Size V.S. Coverage}
As a shared data library, we expect CDS to collect almost all the system related data, the most popular user-generated data and the zero-filled data block.
With CDS as a public data reference, each VM's snapshot backup has no need to store its own copies for data that can be found in CDS, instead they can just
store a reference.

The more data we put into CDS, the closer we come to complete deduplication. But in reality we are facing a list of restrictions such like [blabla].
We borrowed the idea of web caching[refs here] to analysis the size of CDS vesus its effectiveness on data reduction.

We let $M$ be the number of nodes in a cluster, $N$ be the number of VMs hosted per node,
and let $D$ denotes the average size of one VM's snapshot backups (after level 1 and 2 reduction, which is near 40 GB in our production environment).
And let $C_0$, $C_s$ and $C_u$ denote the size of zero-filled block, system-related data and user-generated data in CDS,
then the total size of CDS is represented by $C=C_0+C_s+C_u$.
We let $S_0$, $S_s$ and $S_u$ denote the corresponding data coverage ratio (with respect to $D$) by $C_0$, $C_s$ and $C_u$,
so the total space saving ratio is $S=S_0+S_s+S_u$.

Zero-filled block at maximum length is the first item we need to put into CDS. This is the one and only special data block, 
and because CDS only stores unique data blocks, $C_0$ is almost equal to 0. In practice we found zero-filled blocks account
for near 20\% of total data, so $S_0$ is 20\%.

The rarely modified system-related data are the second to be added to CDS. We have thousands of VMs in each cluster, 
but all these VMs fall into only a few OS types, using a limited selection of software, therefore data duplication in this category
is highly noticable in a global block counting. In particular, most of such data already exist in the VM base images. 
We analyzed VMs running various services from 7 mainstream OS types in our cluster, and found close to 50GB of such data in total. 
Because system-related data don't change frequently, it's safe to expect that for 20 OS variations 
and software updates in 2 years, $C_s$ will not grow to exceed 200 GB.
For each VM, from 2 to 20 GB of data can be reduced in this way, depends on the OS and service type, so on average we estimate
$S_s$ would be no less than 20\%.

The rest of data are user-generated, the total size of such data can be written as $D_u=D-S_0-S_s$, which is 60\% of $D$. 
It follows a zipf-like distribution with $\alpha$ between 0.65 to 0.7. 
Let $T_u$ be the total size of unique data blocks in $D_u$,
in practice we notice that complete deduplication for such data always result in a 2:1 reduction ratio,
, so $T_u=D_u/2$.
Since we collect the most popular user-generated data into CDS, the coverage of CDS on user-generated data
is $(C_u/T_u)^{1-\alpha}$.

The following table lists CDS coverage on user-generated data under different $\alpha$ and $C_u/T_u$.

@@

It's obviously that at least 30\% of user-generated data can be reduced in this way, using about 0.02 of $T_u$, or 0.01 of $D_u$.
The size of user-generated data in CDS would be 0.006$D$, with coverage $S_u=0.18D$.

Thus the estimation of CDS coverage is $S=S_0+S_s+S_u=0.58$, with the size of CDS to be no more than $(200 + 0.006*D*N)$ GB.
In a VM cluster of 100 machines, with 25 VMs on each physical machine, the size of CDS is 800 GB in total, or 8 GB per machine.
The total size of CDS index would be 8 GB, which will cost 80 MB memory on each machine.

>>>>

Overall, our contributions are:
\begin{enumerate}
\item {\em PDS. parallelism, localized}
We designed a highly efficient global deduplication scheme using common data between VMs as the primary
deduplication target, which adopts to our resource-constrained production environment very well. We developed a fast
approxmation algorithm to extract the commonly used data from VM snapshot backups.

@@

\item {\em Fault-tolerant.}
Golbal deduplication brings data sharing across different users VM snapshots, therefore a single failure
on some data can break many VMs' snapshots. We avoid such all-to-all data dependency by restricting
the cross VM data sharing with the PDS. By taking extra care of protecting PDS data, 
the overall degree of fault tolerancy is guaranteed.

@@

% Figure~\ref{fig:arch} shows the architecture view of our snapshot service
% at each node. The snapshot broker provides the functional interface for  snapshot backup, access, and deletion.
% The inner-VM  deduplication is conducted by the broker to access meta data in the snapshot data store
% and we discuss this in details in Section~\ref{sect:innerVM}.
% The cross-VM deduplication is conducted by the broker to access 
% a common data set (PDS) (will discuss in Section~\ref{sect:crossVM},
% whose block hash index is stored in a distributed memory cache. 
% \begin{figure}[htbp]
%   \centering
%   %\epsfig{file=images/arch.eps, height=2in, width=2.66in}
%   \epsfig{file=images/arch.eps, width=3.9in}
%   \caption{Snapshot backup architecture of each node.}
%   \label{fig:arch}
% \end{figure}
% The snapshot store 
% supports data access operations such as \emph{get}, \emph{put} and \emph{delete}.
% Other operations include data block traverse and resource usage report.
% The snapshot data does not need to be
% co-located with VM instances, and in fact they can even live in a different cluster to improve the 
% data reliability: when one cluster is not available, we are still able to restore its VMs from another cluster which
% holds its snapshot data. 
% %to reduce the impact to our users.
% %Get interface accepts a piece of data, write it to the underline data file, and return
% %a reference to the caller. This reference then can be used in the put interface to
% %retrive or delete the data, thus the caller of put interface must preserve the
% %data reference for future use. 
% %In addition to above standand data access operations, the snapshot service also supports
% %\emph{scan} and \emph{quota} methods. Scan allows us to traverse all the data blocks of each VM
% %in the snapshot store, and quota is used to acknowledge user how much space he
% %has actually used.

@@

% \subsection{Block Storage Device}
\subsection{PDS}
Although locality based data reduction can remove most of the inner-VM duplications,
it's not able to solve the data duplication cross VMs. Different VMs still share large amount
of common data such that their snapshot backups would have a lot of data in common.
Our observations on Aliyun's real VM user data reveals several major sources of cross-VM data duplication:
  \begin{enumerate}
  \item \emph{System-related data}: These data are generated by public third-parties, they are copied/downloaded/installed into VM disks either automatically or manually. Once installed, operations on such data are mostly read only until software updates arrive. For example, data of operating systems, some widely-used software such as Apache and MySQL, and their documations all fall into this category.
  \item \emph{User-generated data}: These data are generated by individual user's behavior, either directly or indirectly. They are much less duplicated than the system-related data, but the zipf-like distribution indicates that a small amount of data in this category could represent most of data duplications.
  \item \emph{Zero-filled data}: Like previous studies [refs here], we've observed that zero-filled data exist pervasively at system wide. They are almost like the spaces and commas in text articles. Under content-based chunking, they were distilled as zero-filled blocks with the maximum allowed length. 
%Zero-filled blocks are mostly generated by user behavior to preserve the space for future data, thus we treat them as a special case of user-generated data.
  \end{enumerate}

Without eliminating the data duplication between VM snapshot backups, storage space is severely wasted when the number of VMs increases.
To address this problem, we developed a technique called \emph{Common Data Set (PDS)} to eliminate data duplication for all three situations above. 
PDS is a public data library that shared by all VM snapshot backups in the cluster, 
which consists of the most popular data blocks in VM snapshot backups. It is generated, 
managed and accessed in an uniform manner by all VMs such that [write some fancy system characteristic stuff here].

\subsection{PDS Size V.S. Coverage}
As a shared data library, we expect PDS to collect almost all the system related data, the most popular user-generated data and the zero-filled data block.
With PDS as a public data reference, each VM's snapshot backup has no need to store its own copies for data that can be found in PDS, instead they can just
store a reference.

The more data we put into PDS, the closer we come to complete deduplication. But in reality we are facing a list of restrictions such like [blabla].
We borrowed the idea of web caching[refs here] to analysis the size of PDS vesus its effectiveness on data reduction.

We let $M$ be the number of nodes in a cluster, $N$ be the number of VMs hosted per node,
and let $D$ denotes the average size of one VM's snapshot backups (after level 1 and 2 reduction, which is near 40 GB in our production environment).
And let $C_0$, $C_s$ and $C_u$ denote the size of zero-filled block, system-related data and user-generated data in PDS,
then the total size of PDS is represented by $C=C_0+C_s+C_u$.
We let $S_0$, $S_s$ and $S_u$ denote the corresponding data coverage ratio (with respect to $D$) by $C_0$, $C_s$ and $C_u$,
so the total space saving ratio is $S=S_0+S_s+S_u$.

Zero-filled block at maximum length is the first item we need to put into PDS. This is the one and only special data block, 
and because PDS only stores unique data blocks, $C_0$ is almost equal to 0. In practice we found zero-filled blocks account
for near 20\% of total data, so $S_0$ is 20\%.

The rarely modified system-related data are the second to be added to PDS. We have thousands of VMs in each cluster, 
but all these VMs fall into only a few OS types, using a limited selection of software, therefore data duplication in this category
is highly noticable in a global block counting. In particular, most of such data already exist in the VM base images. 
We analyzed VMs running various services from 7 mainstream OS types in our cluster, and found close to 50GB of such data in total. 
Because system-related data don't change frequently, it's safe to expect that for 20 OS variations 
and software updates in 2 years, $C_s$ will not grow to exceed 200 GB.
For each VM, from 2 to 20 GB of data can be reduced in this way, depends on the OS and service type, so on average we estimate
$S_s$ would be no less than 20\%.

The rest of data are user-generated, the total size of such data can be written as $D_u=D-S_0-S_s$, which is 60\% of $D$. 
It follows a zipf-like distribution with $\alpha$ between 0.65 to 0.7. 
Let $T_u$ be the total size of unique data blocks in $D_u$,
in practice we notice that complete deduplication for such data always result in a 2:1 reduction ratio,
, so $T_u=D_u/2$.
Since we collect the most popular user-generated data into PDS, the coverage of PDS on user-generated data
is $(C_u/T_u)^{1-\alpha}$.

The following table lists PDS coverage on user-generated data under different $\alpha$ and $C_u/T_u$.

@@

It's obviously that at least 30\% of user-generated data can be reduced in this way, using about 0.02 of $T_u$, or 0.01 of $D_u$.
The size of user-generated data in PDS would be 0.006$D$, with coverage $S_u=0.18D$.

Thus the estimation of PDS coverage is $S=S_0+S_s+S_u=0.58$, with the size of PDS to be no more than $(200 + 0.006*D*N)$ GB.
In a VM cluster of 100 machines, with 25 VMs on each physical machine, the size of PDS is 800 GB in total, or 8 GB per machine.
The total size of PDS index would be 8 GB, which will cost 80 MB memory on each machine.
